To Batch or Not to Batch: Sequential vs. Batched Testing Strategies in the ED

### 2023-11-08 manuscript/manuscript.Rmd
working on manuscript

In this section, we examine the causal effects of provider batching tendency. We first present the effects on our main outcomes of length of stay, 72hr readmission, and healthcare resource utilization (number of tests).We find statistically and economically significant positive effects of assignment to a high batching physician on every outcome (Table 3). Estimated coefficients are scaled by the difference in leniency going from the nineti- eth to tenth percentile in physician leniency—equal to 11.6 pp—for interpretability. Assignment to a physician in the top prescribing decile (relative to one in the bottom decile) is associated with a 0.24 pp increase in the probability of long-term prescrip- tion opioid use—a 4.1 percent increase. The effect on proxies for opioid-seeking behavior is 0.50 pp—a 3.4 percent increase. Effects on development of opioid use disorder and on opioid overdose mortality are 2.1 percent and 8.8 percent, respec- tively. These findings highlight the substantial role ED physicians can play in put- ting patients on a path of long-term opioid use and dependence.

>>>>

In this section, we examine the causal effects of provider batching tendency. One thing that we note in the

We first present the effects on our main outcomes of length of stay, 72hr readmission, and healthcare resource utilization (number of tests). We find statistically and economically significant positive effects of assignment to a high batching physician on every outcome (Table 3). Estimated coefficients are scaled by the difference in leniency going from the ninetieth to tenth percentile in physician leniency—equal to 11.6 pp—for interpretability. Assignment to a physician in the top prescribing decile (relative to one in the bottom decile) is associated with a 0.24 pp increase in the probability of long-term prescription opioid use— a 4.1 percent increase. The effect on proxies for opioid-seeking behavior is 0.50 pp—a 3.4 percent increase. Effects on development of opioid use disorder and on opioid overdose mortality are 2.1 percent and 8.8 percent, respec- tively. These findings highlight the substantial role ED physicians can play in putting patients on a path of long-term opioid use and dependence.

\begin{table}[htbp]
\centering
\caption{Reduced Form Regression Results}
\label{tab:reducedform}
\begin{tabular}{p{6cm}ccc}
\toprule
& \multicolumn{1}{c}{\textbf{Log ED LOS}}
& \multicolumn{1}{c}{\textbf{Number of Tests}}
& \multicolumn{1}{c}{\textbf{72hr Return}} \\
& \multicolumn{1}{c}{(1)}
& \multicolumn{1}{c}{(2)}
& \multicolumn{1}{c}{(3)} \\
\midrule
\textit{Model 1} & & & \\
Batch Tendency      & $0.905^{***}$ & $1.715^{***}$ & $-0.025$  \\
                    & $(0.308)$  & $(0.329)$     & $(0.019)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
\midrule
\midrule
\textit{Model 2} & & & \\
Batch Tendency      & $-0.103$      & $0.120^{***}$ & $0.020$   \\
                    & $(0.462)$     & $(0.046)$     & $(0.021)$ \\
                    &     &     &  \\
Testing Inclination    & $0.608^{**}$  & $0.963^{***}$ & $-0.027^{**}$ \\
                    & $(0.242)$  & $(0.021)$     & $(0.011)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the reduced form are for the baseline sample described in the text. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

### 2023-11-08 manuscript/manuscript.tex
working on manuscript

In this section, we examine the causal effects of provider batching
tendency. We first present the effects on our main outcomes of length of
stay, 72hr readmission, and healthcare resource utilization (number of
tests).

>>>>

In this section, we examine the causal effects of provider batching
tendency. We first present the effects on our main outcomes of length of
stay, 72hr readmission, and healthcare resource utilization (number of
tests). We find statistically and economically significant positive
effects of assignment to a high batching physician on every outcome
(Table 3). Estimated coefficients are scaled by the difference in
leniency going from the ninetieth to tenth percentile in physician
leniency---equal to 11.6 pp---for interpretability. Assignment to a
physician in the top prescribing decile (relative to one in the bottom
decile) is associated with a 0.24 pp increase in the probability of
long-term prescription opioid use--- a 4.1 percent increase. The effect
on proxies for opioid-seeking behavior is 0.50 pp---a 3.4 percent
increase. Effects on development of opioid use disorder and on opioid
overdose mortality are 2.1 percent and 8.8 percent, respec- tively.
These findings highlight the substantial role ED physicians can play in
putting patients on a path of long-term opioid use and dependence.

\begin{table}[htbp]
\centering
\caption{Reduced Form Regression Results}
\label{tab:reducedform}
\begin{tabular}{p{7cm}ccc}
\toprule
& \multicolumn{1}{c}{\textbf{Log ED LOS}}
& \multicolumn{1}{c}{\textbf{Number of Tests}}
& \multicolumn{1}{c}{\textbf{72hr Return}} \\
& \multicolumn{1}{c}{(1)}
& \multicolumn{1}{c}{(2)}
& \multicolumn{1}{c}{(3)} \\
\midrule
\textit{Model 1} & & & \\
Batch Tendency      & $0.905^{***}$ & $1.715^{***}$ & $-0.025$  \\
                    & $(0.308)$  & $(0.329)$     & $(0.019)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
\midrule
\midrule
\textit{Model 2} & & & \\
Batch Tendency      & $-0.103$      & $0.120^{***}$ & $0.020$   \\
                    & $(0.462)$     & $(0.046)$     & $(0.021)$ \\
                    &     &     &  \\
Testing Inclination    & $0.608^{**}$  & $0.963^{***}$ & $-0.027^{**}$ \\
                    & $(0.242)$  & $(0.021)$     & $(0.011)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the first stage for the baseline sample described in the text. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Column 3 corresponds to the baseline controls. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}
